---
title: "Lab No. 1"
author: "Elizabeth Walter"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align = 'center')
```

``` {r, warning = FALSE, message = FALSE}
library(tidyverse)
library(dslabs)
data(murders)
```

## 1. What will this conditional expression return and why?

It will return "Not all positives" because one of the elements in the vector x is negative (-3).
```{r q1}
x <- c(1,2,-3,4)

if(all(x>0)){
  print("All Postives")
} else{
  print("Not all positives")
}
```
\newpage
## 2. Which of the following expressions is always FALSE when at least one entry of a logical vector x is TRUE?

a. all(!x) - The expression is looking that every element in the vector satisfies the condition, and the condition is not (indicated by the !) x, or FALSE. all(x)  needs all entries to be TRUE to return TRUE (and therefore doesn't ALWAYS return FALSE when 1 _or more_ entries = TRUE), and any() requires at minimum only one element satisfies the condition and return TRUE.
```{r q2}
x <- c(TRUE, FALSE, FALSE)
all(x)
any(x)
any(!x)
all(!x)
```
\newpage
## 3. The function nchar tells you how many characters long a character vector is. Write a line of code that assigns to the object new_names the state abbreviation when the state name is longer than 8 characters.
```{r q3}
state <- murders$state
abbrev <- murders$abb
new_names <- abbrev[nchar(state) > 8]
new_names
```
\newpage
## 4. Create a function sum_n that for any given value, say n, computes the sum of the integers from 1 to n (inclusive). Use the function to determine the sum of integers from 1 to 5,000.

```{r q4}
sum_n <- function(n){
   x <- 1:n
   sum(x)
}
sum_n(5000)
```
\newpage
## 5. Create a function altman_plot that takes two arguments, x and y, and plots the difference against the sum. Use it to make an altman plot of x <- c(5,7,9) and y <- c(10,11,12). When your function creates the plot, it will output automatically in your Rmarkdown knitted .pdf.

```{r q5}
altman_plot <- function(x,y){
  a <- x + y
  b <- y - x
  plot(a, b)
}

x <- c(5,7,9)
y <- c(10,11,12)

altman_plot(x,y)
```

\newpage
## 6. After running the code below, what is the value of x and why?
The value of x is 3 because it was assigned that value outside of the function. Inside of the function x <- 5 is a local assignment, so it will not change the value of x outside of the function.
```{r q6}
x <- 3
my_func <- function(y){
  x <- 5
  y+5
}
x
```

\newpage
## 7. Write a function compute_s_n that for any given n computes the sum Sn = 1^2 + 2^2 + 3^2 + ... n^2 . Report the value of the sum when n = 10.

```{r q7}
compute_s_n <- function(n){
  x <- 1:n
  y <- x*x
  sum(y)
}

compute_s_n(10)
```
\newpage
## 8. Define an empty numerical vector s_n of size 25 using s_n <- vector("numeric", 25) and store in the results of S1, S2,...,S25 using a for-loop.
```{r q8}
s_n <- vector("numeric", 25)
for(n in 1:25){
  s_n[n] <- compute_s_n(n)
}
s_n
```

\newpage
## 9. Repeat exercise 8, but this time use sapply.
```{r q9}
n <- 1:25
s_n <- sapply(n, compute_s_n)
s_n
```

\newpage
## 10. Repeat exercise 8, but this time use map_dbl.
```{r q10}
n <- 1:25
s_n <- map_dbl(n, compute_s_n)
s_n
```

\newpage
## 11. Plot Sn versus n. Use points defined by n = 1,...,25 .
```{r q11}
plot(n, s_n)
```

\newpage
## 12. Confirm that the formula for this sum is Sn = n(n+1)(2n+1)/6.
```{r q12}
form <- function(x){
  y <- n*(n+1)*(2*n+1)/6
}
  
for(n in 1:25){
  s_n <- form(n)
}
s_n
```
