---
title: "Lab 3"
author: "Elizabeth Walter"
date: "19 September 2021"
output:
  pdf_document: default
  html_document: default
---


# Essential pandemic construction

The New York City Department of Buildings (DOB) maintains [a list of construction sites](https://www1.nyc.gov/assets/buildings/html/essential-active-construction.html) that have been categorized as "essential" during the city's shelter-in-place pandemic order.

## Load and clean data

First we load and clean the data.

```{r load-clean-data, warning=FALSE, message=FALSE}
# You'll only need the tidyverse library for this exercise
library(tidyverse)

# Load original data; note that this needs to point to the file you downloaded.
essential_raw <- read_csv("EssentialConstruction.csv")

# Clean the data a little
# Some of the borough names are in ALL CAPS, so we use str_to_title() to convert
# everything in the column to title case.
# We also make BOROUGH and CATEGORY factors (or categorical variables)
essential <- essential_raw %>% 
  mutate(BOROUGH = str_to_title(BOROUGH),
         BOROUGH = factor(BOROUGH),
         CATEGORY = factor(CATEGORY))
```

## Approved projects by borough

Right now there's a row for each approved construction site. We need to condense that down to get counts of construction sites by different variables. We can do this by using `group_by()` and `summarize()`

```{r summarize-data-borough}
essential_by_borough <- essential %>% 
  group_by(BOROUGH) %>% 
  summarize(total = n()) %>% 
  mutate(proportion = total / sum(total))
```

```{r plot-borough-summary}
# Add plot with geom_col() here
apprvd_by_bor <- ggplot(data = essential_by_borough, aes(x = BOROUGH, y = total)) +
  geom_col() +
  labs(x = 'Borough', y= 'Total',
       title = 'Approved Projects by Borough')
apprvd_by_bor
```


## Approved projects by category

```{r summarize-data-category}
# Create a summarized dataset of projects by category
essential_by_categ <- essential %>%
  group_by(CATEGORY) %>%
  summarize(total = n()) %>% 
  mutate(proportion = total / sum(total))
```

```{r plot-category-summary}
# Add a lollipop chart here
apprvd_by_categ <- ggplot(data = essential_by_categ, aes(x = CATEGORY, y = proportion)) +
  geom_point(color = 'blue') + 
  geom_segment(aes(x= CATEGORY, xend = CATEGORY, y = 0, yend = proportion)) +
  labs(title = "Lollipop Graph of Projects by Category") +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, vjust = .6))
apprvd_by_categ
```
Info/code for lollipop graphs from: https://www.r-graph-gallery.com/300-basic-lollipop-plot.html and
https://www.r-graph-gallery.com/301-custom-lollipop-chart.html#general 

## Approved projects across borough and category

```{r summarize-data-heatmap}
# Create a summarized dataset of projects by both borough and category

essential_bor_categ <- essential %>%
  select(BOROUGH, CATEGORY) %>%
  group_by(BOROUGH, CATEGORY) %>% 
  summarize(tot_cat= n()) %>% 
  group_by(BOROUGH) %>%
  mutate(tot_bor = sum(tot_cat),
         cat_prop_of_bor = tot_cat / tot_bor)
essential_bor_categ
```

```{r plot-heatmap}
# Add a heatmap here with geom_tile()
bor_categ_heatmap <- ggplot(data = essential_bor_categ, aes(x = BOROUGH, y = CATEGORY, 
                                                            fill = cat_prop_of_bor)) +
  geom_tile() +
  guides(fill = guide_colorbar(title = "Proportion")) +
  labs(title = "Heatmap of Approved Projects by Borough & Category")
bor_categ_heatmap
```
Additional information/code about heatmaps from: https://r-charts.com/correlation/heat-map-ggplot2/
