---
title: "Lab No. 6"
author: "Elizabeth Walter"
output: pdf_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align = 'center')
```

```{r, warning=FALSE, message=FALSE}
library(tidyverse)  # For ggplot, dplyr, and friends
library(patchwork)  # For combining ggplot plots
library(GGally)     # For scatterplot matrices
library(broom)      # For converting model objects to data frames

ameslist <- read.csv("C:/Users/walte/Desktop/MSU SSQDA/SSC 442/Data/ames.csv",
                 header = TRUE,
                 sep = ",")
```
## 1. Prune the data to all of the variables that are type = int about which you have some reasonable intuition for what they mean. This must include the variable SalePrice. Save this new dataset as Ames. Produce documentation for this object in the form of a .txt file. This must describe each of the preserved variables, the values it can take (e.g., can it be negative?) and your interpretation of the variable.

```{r q1}
# list class of each var in ameslist & get indexes of vars in ameslist where class = int
b <- which(sapply(ameslist, class) %in% c('integer'))
length(b) # expect 38 variables

# get list of vars in ameslist of type int
names(ameslist[b])

# get new dataframe of only variables of type int
Ames <- ameslist[names(ameslist[b])]

q1 = read.csv("C:/Users/walte/Desktop/MSU SSQDA/SSC 442/Labs/Lab 6/Q1 documentation", 
              header=TRUE, sep = ",")
q1
```


\newpage
## 2. Produce a scatterplot matrix which includes 12 of the variables that are type = int in the data set. Choose those that you believe are likely to be correlated with SalePrice.

```{r q2}
price_corrs <- Ames %>%
  select(LotArea, OverallQual, OverallCond, YearBuilt, YearRemodAdd, GrLivArea, 
         FullBath, HalfBath, BedroomAbvGr, TotRmsAbvGrd, YrSold, GarageArea, SalePrice)

ggpairs(price_corrs)
```

\newpage
## 3. Compute a matrix of correlations between these variables using the function cor(). Does this match your prior beliefs? Briefly discuss the correlation between the miscellaneous variables and SalePrice.

All of the variables that are positively correlated with SalePrice match my prior beliefs of what their direction of correlation would be. OverallQual has the strongest correlation to SalePrice (both negative and positive) of all the variables chosen, with a positive correlation of r = 0.79. I believed that there would be a high positive correlation between those two variables. GrLivArea also has a strong positive correlation (r = 0.71), which I expected. I thought LotArea and SalePrice would have a stronger positive correlation than r = 0.26. GarageArea and SalePrice have a stronger correlation than I expected (r = 0.62), but I expected it to be moderate and positive. I am surprised that BedroomAbvGr and SalePrice are not strongly correlated (r = 0.17), and more so that it is weaker than the correlation between HalfBath and SalePrice (r = 0.28). I would have expected it to be similar to, if not greater than, the correlation between SalePrice and FullBath, which I also expected to be a strong correlation, but is slightly weaker than expected (r = 0.56). The correlation between YearBuilt and SalePrice is similar to what I expected (r = 0.52).

The only negatively correlated variables are also pretty weakly correlated. I thought that YrSold might be more correlated with SalePrice, but it appears to have no correlation (r = - 0.03), and I expected OverallCond and SalePrice to have a positive and stronger correlation than the very weak, negative correlation they have (r = -0.08).
```{r q3}
ames_corr <- Ames %>%
  select(LotArea, OverallQual, OverallCond, YearBuilt, YearRemodAdd, GrLivArea, 
         FullBath, HalfBath, BedroomAbvGr, TotRmsAbvGrd, YrSold, GarageArea, SalePrice) %>%
  cor()

ames_corr
```

\newpage
## 4. Produce a scatterplot between SalePrice and GrLivArea. Run a linear model using lm() to explore the relationship. Finally, use the abline() function to plot the relationship that you’ve found in the simple linear regression.

```{r q4}
ggplot(Ames, aes(x = GrLivArea, y = SalePrice)) +
  geom_point()
```

```{r }
price_livarea_reg <- lm(SalePrice ~ GrLivArea, data = Ames)
tidy(price_livarea_reg, conf.int = TRUE)
```

```{r }
ggplot(Ames, aes(x = GrLivArea, y = SalePrice)) +
  geom_point() +
  geom_abline(intercept = 18569.0, slope = 107.1)
```

### 4b. What is the largest outlier that is above the regression line? Produce the other information about this house.

Largest outlier ABOVE the regression line is a house with GrLivArea of 4316 and SalePrice of $755,000
```{r }
out_above <- Ames[which.max(Ames$SalePrice),]
out_above
```
Largest outlier BELOW the regression line (also house with greatest distance from regression line) is a house with GrLivArea of 5642 and SalePrice of $160,000.
```{r }
out_below <- Ames[which.max(Ames$GrLivArea),]
out_below
```
